A note on the instability and degeneracy of deep learning models

### 2017-09-27 writing/paper-note/appendices/appendix-note.Rmd
some changes in the note from nordman

**Proof of Proposition \@ref(prp:degenFSFS).** where $|\mathcal{X}|<\infty$ holds in the FSFS model. We may suppose $|\mathcal{X}|>1$ (i.e., $\mathcal{X}^N$ has more than one outcome) because otherwise the model is trivially degenerate for all $N \geq 1$. Fix $0 < \epsilon < 1$. Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta_N}$, so $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \ge P_{\boldsymbol \theta_N}(\boldsymbol x_{max}) > 0$. If $\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}$, then by definition $P_{\boldsymbol \theta_N}(\boldsymbol x) \le [P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}$ holds so that
\begin{align*}
1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})
& = \sum\limits_{\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}}P_{\boldsymbol \theta_N}(\boldsymbol x) \\
& \le (|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}.
\end{align*}
From the lower bound on $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$ and the upper bound on $1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$, it follows that
\begin{align*}
\frac{1}{N}\log\left[\frac{P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}{1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}\right] & \ge \frac{1}{N} \log\left[\frac{P_{\boldsymbol\theta_N}(\boldsymbol x_{max})}{(|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{N} \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] - \log |\mathcal{X}| \rightarrow \infty
\end{align*}
as $N \rightarrow \infty$ by the definition of an unstable FSFS model (cf. Definition \@ref(def:instabFSFS)). Consequently, $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \rightarrow 1$ as $N \rightarrow \infty$ as claimed. \hfill $\Box$

>>>>

**Proof of Proposition \@ref(prp:degenFSFS).** Where $|\mathcal{X}|<\infty$ holds in the FSFS model, we may suppose $|\mathcal{X}|>1$ (i.e., $\mathcal{X}^N$ has more than one outcome) because otherwise the model is trivially degenerate for all $N \geq 1$. Fix $0 < \epsilon < 1$. Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta_N}$, so $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \ge P_{\boldsymbol \theta_N}(\boldsymbol x_{max}) > 0$. If $\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}$, then by definition $P_{\boldsymbol \theta_N}(\boldsymbol x) \le [P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}$ holds so that
\begin{align*}
1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})
& = \sum\limits_{\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}}P_{\boldsymbol \theta_N}(\boldsymbol x) \\
& \le (|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}.
\end{align*}
From the lower bound on $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$ and the upper bound on $1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$, it follows that
\begin{align*}
\frac{1}{N}\log\left[\frac{P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}{1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}\right] & \ge \frac{1}{N} \log\left[\frac{P_{\boldsymbol\theta_N}(\boldsymbol x_{max})}{(|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{N} \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] - \log |\mathcal{X}| \rightarrow \infty
\end{align*}
as $N \rightarrow \infty$ by the definition of an S-unstable FSFS model (cf. Definition \@ref(def:instabFSFS)). Consequently, $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \rightarrow 1$ as $N \rightarrow \infty$ as claimed. \hfill $\Box$
